---
title: "02_animated_ggmap_drowned"
author: "Duc-Quang Nguyen"
date: "16 June 2016"
output: html_document
---

## Data

* columns *CartoDB_Cause_of_death* has 7 different type of death/disappearance
* 

```{r setup, include=FALSE}
library(dplyr)
library(magrittr)
library(tidyr)
library(countrycode)
library(swiMap)
require(rgdal)
library(rgeos)
library(maptools)
require(ggplot2)
library(swiTheme)
library(gganimate)
library(extrafont)

translation.file <- "input/Death in the Mediterranean over the last year - Sheet1.csv"
exludeLangs <- c("EN", "DE", "IT", 'PT', "FR", "ES")

```

```{r load data}
data.read <- read.csv("data/Events during which someone died trying to reach or stay in Europe - Events.csv", stringsAsFactors = F, check.names = F)

# format date
data.read$date <- as.Date(data.read$date)

# drop columns
data <- data.read %>% select(-quarter, -`Intent of going to Eur: 1(yes) 0(not confirmed)`, -`Somme Dedoublement`, -`name`)

# recreate lon & lat columns (they lon lat have different formatting)
latlon <- data %>% select(`latitude, longitude`) %>% unlist(use.names = F)
latlon <- do.call(rbind, sapply(strsplit(latlon, ", "), function(i) {
  if(length(i) == 0) {
    c(NA, NA)
  } else {
    as.numeric(i)
  }
}))
colnames(latlon) <- c("lat", "lon")
data <- cbind(data, latlon)


txt <- read.csv(translation.file, row.names = 1, stringsAsFactors = F)
# discard incomplete translations
cidx <- unique(which(txt =="" | is.na(txt), T)[,2])
if(length(cidx > 0)) {
  warning(paste(colnames(txt)[cidx], collapse = "\t"), " languages will be discarded!", "\n")
  txt <- txt[,-cidx, drop = F]
}
colnames(txt)


```

```{r prepare data for cartodb, echo = F}
# date.from <- as.Date("2015-06-20")
# cdb <- filter(data, date >= date.from, cause_of_death == "drowned")
# # ugly hack to remove  events looking weird not in the sea at all!
# cdb <- filter(cdb, !Event_id %in% c(34025, 33990, 33864, 33842, 33908), !is.na(lon))
# 
# cdb %<>% select(-dataset, -CartoDB_Cause_of_death, -cause_of_death, -Event_id,
#   -description, -location, -latitude, -longitude, -source, -source_url, -`route (Frontex)`)
# 
# write.csv(cdb, file = "input/lastYear_drownedMigrants4CDB.csv", row.names = F)

```

```{r map}
date.from <- as.Date("2015-06-20")
txt.xpos <- 37
txt.ypos <- 55


#fontSize <- "0.9em"

dd <- filter(data, date >= date.from, cause_of_death == "drowned")
# ugly hack to remove  events looking weird not in the sea at all!
dd <- filter(dd, !Event_id %in% c(34025, 33990, 33864, 33842, 33908))
dates <- seq(date.from, max(dd$date), 1)

# drop columns
dd %<>% select(-cause_of_death, -CartoDB_Cause_of_death, -description, -source, -source_url, -latitude, -longitude)

## Map

world <- map_data("world")
# filter
reg2cont <- unique(world$region)
names(reg2cont) <- countrycode(reg2cont, "country.name", "continent")
world$continent <- names(reg2cont)[match(world$region, reg2cont)]
world <- world[world$region != "Antarctica",]
world <- world %>% 
  filter(region !=  "Antarctica", !continent %in% c('Americas', 'Oceania'))


wm <- ggplot() + 
  geom_polygon(data = world, aes(x = long, y = lat, group = group), 
    size = 0.2, fill = "#d4c3aa", colour = "#aa8959") + 
  swi_theme(y_gridlines = F, base_size = 16) + 
  theme(
    panel.background = element_rect(fill = '#f7f5ed', size = 0),
    panel.grid = element_blank(), 
    axis.ticks = element_blank(), 
    axis.title = element_blank(), 
    axis.text = element_blank(),
    legend.position = "none",
    plot.title = element_text(family="Open Sans Condensed", size = 23, margin=margin(b=10), face = "bold"),
    plot.caption=element_text(size=12, hjust=0, margin=margin(t=10)),
    plot.margin=unit(rep(0, 4), "cm")
  )

 # With nice orthographic projection
wm <- wm + coord_map("orthographic",  orientation = c(30, 12, 0), xlim = c(-11, 43), ylim = c(24, 60))
dd <- arrange(dd, date)

dd$cumDM <- cumsum(dd$dead_and_missing)
dates <- seq(min(dd$date), max(dd$date), by = "week")
dd$startWDate <- dates[findInterval(dd$date, dates)]


counter <- dd %>% group_by(startWDate) %>% summarise(dm = sum(dead_and_missing)) %>% ungroup()
counter$cumsumDM <- cumsum(counter$dm)
counter$x <- txt.xpos
counter$y <- txt.ypos

# duplicate the last date
counter <- counter[c(1:nrow(counter), rep(nrow(counter), 3)),]
# hack to add a date£
counter[nrow(counter),'startWDate'] <- counter[nrow(counter),'startWDate'] + 7
#counter$label <- paste0("dead or missing\n", counter$cumsumDM)

langs <- colnames(txt)[!colnames(txt) %in% exludeLangs]


sapply(langs, function(lang) {
  cat("\n", lang, "\n")
  if(lang == "JA") {
    txt.xpos <- txt.xpos - 6 
  }
  output <- paste0("migrants_deathAtSea_", lang, ".gif")
  
  p <- wm + geom_point(
    data = dd, aes(x = lon, y = lat, size = dead_and_missing, frame = startWDate, cumulative = T), 
    color = "#ab3d3f", alpha = 0.7
  ) + 
  scale_size(range = c(1, 22)) +
    annotate(
      "text", x = txt.xpos, y = txt.ypos, label = paste0( txt['dm', lang], "  "),
      size = 14, color = "#ab3d3f", family = txt['txt.title', lang], hjust = 1, vjust = 0, alpha = 0.8) +
    geom_text(
      data = counter, aes(x = x, y = y, label = cumsumDM, frame = startWDate), 
      size = 27, color = "#ab3d3f", family = txt['txt.number', lang], fontface = "bold", hjust = 0, vjust = 0, alpha = 0.9) +
    labs(
      title = paste0 (txt['main.title', lang],  "\n", txt['main.title2', lang], ":"), 
      caption = txt['footer', lang]
    ) + 
    theme(
        plot.title = element_text(family = txt['font.title', lang], size = 22, margin=margin(b=12), face = "bold")
    )
  if(lang == "JA") {
    p <- p + theme(plot.caption=element_text(family =  txt['font.title', lang]))
  }
  gg_animate(p, output, interval = 0.65, ani.width = 700, ani.height = 640)
})



# p <- wm + geom_point(
#   data = dd, aes(x = lon, y = lat, size = dead_and_missing, frame = startWDate, cumulative = T), 
#   color = "#ab3d3f", alpha = 0.7
#   ) + scale_size(range = c(1, 22)) +
#   annotate("text", x = txt.xpos, y = txt.ypos, label = "dead or missing: ",
#     size = 14, color = "#ab3d3f", family = "OpenSans-CondensedLight", hjust = 1, vjust = 0, alpha = 0.8) +
#   geom_text(data = counter, aes(x = x, y = y, label = cumsumDM, frame = startWDate), 
#     size = 27, color = "#ab3d3f", family = "Open Sans", fontface = "bold", hjust = 0, vjust = 0, alpha = 0.9) +
#   labs(
#     title = paste0 ("Migrants who have drowned over the last year", ": "), 
#     caption = "source: The Migrants' Files | swissinfo.ch"
#   ) +
# 
# gg_animate(p, "migrants_deathAtSea.gif", interval = 0.6, ani.width = 700, ani.height = 600)
